Future challenges for Research Software Engineering

### 2025-05-17 future_challenges_for_RSEng.tex
reformulation

Research Software Engineering is still a developing field,
with increasing integration into various aspects of research.
In this opinion paper, we try to imagine the state of our field
and community in the mid-term future, focusing on future challenges.
We present here the results of a workshop discussion at the
deRSE25 conference in Karlsruhe\cite{Goth2025EndRSEng}, without
claiming to represent the entire deRSE community.
Nevertheless, by disseminating the discussed ideas,
we aim to engage the community in creating a future worth living
and advance the conversation about how Research Software Engineering should, or could, develop.
In the following sections, we consider a number of areas:
the evolution of the RSE definition;
the development of RSE education;
the growth of the RSE communities;
the impact of automation and generative artificial intelligence in RSE work;
the emerging ethical considerations;
the transformations expected in research organisations;
and, finally, if there will be other areas in academia covered by new groups of RTPs (\emph{Research Technical Professional}s).
We conclude the paper with a summary and with our wishes and recommendations.

>>>>

Research Software Engineering is still a developing field,
with increasing integration into various aspects of research.
In this opinion paper, we try to imagine the state of our field
and community in the mid-term future, focusing on future challenges.
We present here the results of a workshop discussion at the
deRSE25 conference in Karlsruhe\cite{Goth2025EndRSEng}, without
claiming to represent the entire deRSE community.
Nevertheless, by disseminating the discussed ideas,
we aim to engage the community in creating a worthwhile future
and to advance the conversation about how Research Software Engineering should, or could, develop.
In the following sections, we consider a number of areas:
the evolution of the RSE definition;
the development of RSE education;
the growth of the RSE communities;
the impact of automation and generative artificial intelligence in RSE work;
the emerging ethical considerations;
the transformations expected in research organisations;
and, finally, if there will be other areas in academia covered by new groups of RTPs (\emph{Research Technical Professional}s).
We conclude the paper with a summary and with our wishes and recommendations.
